---
title: "Lab 02 - Describe Data"
output: html_document
date: "`r Sys.Date()`"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(ggplot2)
library(gridExtra)
```

## 1 Load Data
### 1.1 Load DatasSet
```{r}
df <- read_delim("../../data/king-house-prices.csv.bz2")
```

### 1.1.2a Check rows and columns
```{r}
dim(df)
```
There are 21613 rows and 21 columns

### 1.1.2b Print few lines of data
```{r}
sample_n(df, 10)
```
I've print 10 random rows as example. All these data looks good with correct data type and the values looks reasonable.

```{r}
colSums(is.na(df))
```
There's no null values in this dataset.

## 2. Listing price
### 2.1 Find max and min of listing price
```{r}
range(df$price)
```
The price of home range from 75,000 to 7,700,000. The price of housing was reasonable.

### 2.2 Compute mean and median
```{r}
df %>% 
  summarise(mean = mean(price), median = median(price))
```
The mean listing price was $540088.1 and the median listing price was $450000. The mean value is $9000 higher than median value which means that the distribution is skew to the right with outlier of much expensive houses.

### 2.3 compute standard deviation
```{r}
price <- df$price
sqrt(mean((price-mean(price))^2))
```
The standard deviation of price was $367118.7

### 2.4 compare sd calculation method
```{r}
sd(price)
```
the previous result is around $9 less than sd method. Since sd method uses 1/(N-1) instead of 1/N

### 2.5 plot price in linear scale and log scale
```{r}
l <- ggplot(df, aes(price)) +
  geom_histogram(bins = 33, fill = "pink", color="cyan") 
s <- ggplot(df, aes(price)) +
  geom_histogram(bins = 33, fill = "pink", color="cyan") +
  scale_x_log10()
grid.arrange(l, s, ncol = 2)
```

### 2.6 Comment on histograms
For the linear scale graph on the left, the distribution is heavily right skewed with a really high tail at $8e+06. If we look at the log scale graph on the right, it shows us a normal distribution.

## 3. Sample proportions
### 3.1 Check percentage of cheap and expensive price
```{r}
df %>% 
  summarize(cheap = mean(df$price < 200000), expensive = mean(df$price > 1000000))
```
3.6% of the listing price are cheaper than $200,000 and 6.78% of the listing price are more expensive than $1M

### 3.2 Price less than 1M and overlook waterfront
```{r}
df %>% 
  summarise(mean(price < 1000000 & waterfront))
```
There are 0.268% of listing price that are less than $1M and overlooking waterfront

## 4 Condition and price
### 4.1 Median, mean, and sd of price in each of condition
```{r}
df %>% 
  group_by(condition) %>% 
  summarise(median = median(price), mean = mean(price), sd = sd(price))
```
### 4.2 Make boxplot of relationship
```{r}
ggplot(df, aes(as_factor(condition), price)) +
  geom_boxplot() + 
  scale_y_log10()
```

### 4.3 Comment on lab
For aparment with 1 and 2 index of condition, there wasn't much outlier. The graph demonstrate that the price tends to be higher for better condition apartment. 

