---
title: "INFO370_lab05"
author: "Marco"
date: "`r Sys.Date()`"
output: html_document
---
```{r}
library(tidyverse)
library(ggplot2)
```

## 1
### 1.1
```{r}
dd <- read_delim("../../data/diamonds.csv.bz2")
dim(dd)
sample_n(dd, 5)
```
There are 53940 dinomds

### 1.2
```{r}
table(dd$cut)
```


There are 1610 Fair, 4906 Good, 21551 Ideal, 13791 Premium, and 12082 Very Good.

## 2
### 2.1
```{r}
m <- lm(price ~ carat + cut , data = dd)
summary(m)
```
### 2.2
The reference category would be fair for cut

### 2.3.a
The carat is when the diamond's cut is fair, one additional carat of diamond would be 7871.08 more expensive

### 2.3.b
When the carat of two diamonds remain the same, the Ideal cut diamond would be 1800.92 more expensive than fair cut diamond.

### 2.3.c
```{r}
1800.92-1510.14
```
In average, ideal-cut diamonds are $290.78 more expensive than very good cut diamonds within same carat.

### 2.3.d
```{r}
7871.08-3875.47+1439.08
```
My model predict that the average price for 1ct premium cut diamond is $5434.69

## 3
### 3.1
```{r}
ggplot(sample_n(dd, 1000), aes(y = price, x = carat)) +
  geom_smooth(method = "lm", col = "purple")+
  geom_point(alpha = 0.6, col = "gold2")
```
Our linear model doesn't capture the the relationship between price cut and carat 
### 3.2
```{r}
ggplot(sample_n(dd, 1000), aes(y = log(price), x = carat)) +
  geom_smooth(method = "lm", col = "purple")+
  geom_point(alpha = 0.6, col = "gold2") +
  labs(title = "Log Scale of price")
```

```{r}
ggplot(sample_n(dd, 1000), aes(y = price, x = log(carat))) +
  geom_smooth(method = "lm", col = "purple")+
  geom_point(alpha = 0.6, col = "gold2") +
  labs(title = "Log Scale of carat")
```

```{r}
ggplot(sample_n(dd, 1000), aes(y = log(price), x = log(carat))) +
  geom_smooth(method = "lm", col = "purple")+
  geom_point(alpha = 0.6, col = "gold2") +
  labs(title = "Log Scale of both carat and price")
```
The plot of both variables in log scale looks the best


### 3.3 
```{r}
m2 <- lm(log(price) ~ log(carat) + cut , data = dd)
summary(m2)
```

### 3.4
I got a pretty high R square of 0.9371 which means the model did great on capturing the relationhship between price carat and cut

### 3.5
The intercept coefficient means that when the fair cut diamond is 1 carat $e^0$ the price of it would be $e^{8.2}$.
The log(carat) coefficient means that when you increase the log(carat) of a diamond by one the price of it would be increased by $e^{1.69}$ percent.
The cutGood coefficient means that when you get two diamond of same carat, the good cut one would be 0.163245 percent more expensive than the fair cut one.
The cutIdeal coefficient means that when you get two diamond of same carat, the ideal cut one would be 0.317212 percent more expensive than the fair cut one.
The cutPremium coefficient means that when you get two diamond of same carat, the Premium cut one would be 0.238217 percent more expensive than the fair cut one.
The cutVery Good coefficient means that when you get two diamond of same carat, the Premium cut one would be 0.240753 percent more expensive than the fair cut one.