---
title: "grouped"
output: html_document
date: "2023-11-07"
---

```{r}
library(tidyverse)
titanic <- read_delim("../Data/titanic.csv")
dim(titanic)
```

```{r}
titanic %>% 
  summarize(fare = mean(fare, na.rm=TRUE))
```

```{r}
titanic %>% 
  filter(sex == "female") %>% 
  summarize(fare = mean(fare, na.rm=TRUE))
```

```{r}
titanic %>% 
  summarize(n_distinct(cabin))
```

#### put men and women is different group (seperate data set)
```{r}
titanic %>%
  group_by(sex) %>%
  summarize(fare = mean(fare, na.rm=TRUE))
```

```{r}
titanic %>%
  group_by(cabin) %>%
  summarize(fare = mean(fare, na.rm=TRUE))
```

#### How much for the most expensive cabin and how many people
```{r}
titanic %>%
  group_by(cabin) %>%
  summarize(fare = mean(fare, na.rm=TRUE), n()) %>% 
  arrange(desc(fare))
```

#### What is the biggest cabin
```{r}
titanic %>%
  group_by(cabin) %>%
  summarize(fare = mean(fare, na.rm=TRUE), n = n()) %>% 
  arrange(desc(n))
```


#### Load as fild
```{r}
library(tidyverse)
bn <- read_delim("../Data/babynames.csv.bz2")
dim(bn)
```


#### load packages
```{r}
library(babynames)
dim(babynames)
head(babynames)
```

How many different names b/w 2000 and 2005?
```{r}
babynames %>% 
  filter(between(year, 2000, 2005)) %>% 
  summarize(n = n_distinct(name))
```

How many different names for each of these years
```{r}
babynames %>% 
  group_by(year) %>% 
  filter(between(year, 2000, 2005)) %>% 
  summarize(n = n_distinct(name))
```

How many boys and girls name each year
```{r}
babynames %>%
  group_by(year, sex) %>%
  filter(between(year, 2000, 2005)) %>%
  summarize(n = n_distinct(name))
```

What is the most popular name
```{r}
babynames %>%
  filter(between(year, 2000, 2005)) %>%
  arrange(desc(n)) %>% 
  head(3)
```

Which one is the smallest in a 
```{r}
x <- c(12, 11, 13)
rank(x)
```

```{r}
x[rank(x) == 1]
```
Order in different arrangement
```{r}
babynames %>%
  filter(between(year, 2000, 2005)) %>%
  filter(rank(desc(n)) <= 3)
```
Exercise:

What are the 3 most popular girls/boy is this period

```{r}
babynames %>%
  group_by(sex) %>% 
  filter(between(year, 2000, 2005)) %>%
  filter(rank(desc(n)) <= 3)
```

```{r}
babynames %>%
  group_by(sex) %>%
  filter(between(year, 2000, 2005)) %>%
  filter(rank(desc(n)) <= 3) %>%
  arrange(sex, desc(n))
```
Exercise:

What are the most popular girl/boy names for each year in this period
```{r}
babynames %>%
  group_by(sex, year) %>%
  filter(between(year, 2000, 2005)) %>%
  filter(rank(desc(n)) <= 1) %>%
  arrange(sex, year)
```

How many times the gender were given in these data?
```{r}
babynames %>%
  filter(between(year, 2000, 2005)) %>% 
  group_by(sex) %>% 
  summarize(N = sum(n))
```

How many times the gender were given in these data?
```{r}
babynames %>%
  filter(between(year, 2000, 2005)) %>% 
  group_by(name) %>% 
  summarize(N = sum(n)) %>% 
  arrange(desc(N)) %>% 
  head(6)
```

Find 3 most populat names 2000-2005 for boys and girls
```{r}
babynames %>%
  filter(between(year, 2000, 2005)) %>%
  group_by(name, sex) %>%
  summarize(N = sum(n)) %>%
  group_by(sex) %>%
  filter(rank(desc(N)) <= 3) %>%
  arrange(sex, desc(N)) %>% 
  ungroup()
```

How popular was name "Mei" for girls in 2004?
```{r}
babynames %>% 
  filter(year == 2004,
         sex == "F") %>% 
  mutate(r = rank(desc(n))) %>% 
  filter(name %in% c("Mei", "Madison"))
```

```{r}
n <- c(400, 800, 200, 350)
rank(n)
```

