---
title: "PS4_DataFrame"
output: html_document
date: "2023-11-12"
---


### 1.1 
```{r}
getwd()
```

### 1.2
"/Users/guojiawei/Downloads/INFO 201/PS"

### 1.3
Yes, they are in the same folder. If the working directory is not the same as the folder where my .rmd file is, relative paths might not work as expected because r might not find these files.

### 2.1 
Create a vector “names” with (at least) 5 names in it
```{r}
names <- c("Jacob", "Emma", "Emily", "Noah", "Madison")
```

### 2.2 
Create a vector “math” with grades (0 - 100) in a math course, one grade for each student.
```{r}
math <- sample(100, 5, replace = TRUE)
```

### 2.3 
Create a vector “japanese” with grades for the Japanese course
```{r}
japanese <- sample(100, 5, replace = TRUE)
```

### 2.4 
Create a vector “dance” with grades for a dance class
```{r}
dance <- sample(100, 5, replace = TRUE)
```

### 2.5 
Create a data frame “grades” by combining your names, and math, Japanese and dance grades
```{r}
grades <- data.frame(names, math, japanese, dance)
```

### 2.6 
Compute the number of students in your data!
<!-- get number of students in data -->
```{r}
nrow(grades)
```

### 2.7 
Compute number of courses in data, minus one because names column
<!-- get the number of courses -->
```{r}
ncol(grades) - 1
```

### 2.8 
Print the last two lines of your data
<!-- shows the last two lines of data -->
```{r}
print(tail(grades, 2))
```

### 2.9
Save students’ data into a new .csv file inside of data/ directory as “grades.csv”
```{r}
library(readr)
write_csv(grades, "../Data/grades.csv")
```

### 2.10 
Show that saved grades.csv is there
```{r}
list.files("../Data/")
```

### 3.1
create a variable “col” that contains name of one of the courses–dance.
```{r}
col <- "dance"
```

### 3.2
Print dance grades using the variable “col”
```{r}
grades[, col]
```

### 3.3
Assign name of course-math to “col” and print the best grade in that course!
```{r}
col <- "math"
max(grades[, col])
```

### 3.4
print the name of the course variable, and its average value
```{r}
for (col in names(grades)) {
    if (is.numeric(grades[, col])) {
        cat(col, "course has average grade of:", mean(grades[, col]), "\n")
    }
}
```

### 4.1
GPA = 1/3 (math + Japanese + dance)
Use vectorized operations, Save it as variable ’gpa’ indata frame
```{r}
library(tidyverse)
grades <- grades %>% 
  mutate(GPA = (math+japanese+dance)/3)
grades$GPA
```

### 4.2
Print that line of person with best GPA
```{r}
grades %>% 
  arrange(desc(GPA)) %>% 
  head(1)
```
### 4.3
print only the best student’s name
```{r}
best <- grades %>% 
  filter(GPA == max(GPA))
print(best$names)
```

### 4.4
print the data
```{r}
print(grades)
```
it is correct

### 4.5
Compute variable that is TRUE/FALSE, depending on if the student better at math than in Japanese
Show three lines of your data that includes the new variable.
```{r}
grades <- grades %>%
  mutate(BetterMathThanJapanese = math > japanese)
head(grades, 3)
```

### 4.6
```{r}
sum(grades$BetterMathThanJapanese)
```

### 5.1.1
Load life-expectancy.csv file
```{r}
life_expectancy <- read_delim("../Data/life-expectancy.csv.bz2")
```
### 5.1.2
Number of rows and columns life-expectancy.csv contain?
```{r}
dim(life_expectancy)
```

### 5.1.3
Print its variable names
```{r}
names(life_expectancy)
```

### 5.1.4
```{r}
life_expectancy %>% 
  sample_n(5)
```
It does look good since it produce an organized table with text, number, and missing values. It doesn't look carbled or empty.

### 5.1.5
Each row is representing a country

### 5.1.6
```{r}
for (col in names(life_expectancy)) {
    missing_count <- sum(is.na(life_expectancy[[col]]))
    cat(col, "has", missing_count, "missing values\n")
}
```
#### The name and region variables are good, the le1960, le2019, GDP_PC1960, GDP_PC2019, population1960, population2019 contains some missing values.

### 5.2.1
Create a new variable in the dataset–the growth in life expectancy from 1960 to 2019. Display a few lines of your augmented data to show it worked.
```{r}
life_expectancy <- life_expectancy %>% 
  mutate(growthLE = le2019 - le1960) 
sample_n(life_expectancy, 5)
```
### 5.2.2
What is the average improvement in LE over these years?
```{r}
mean(life_expectancy$growthLE, na.rm = TRUE)
```
#### the average improvement in LE is 18.79684

### 5.2.3
Which country gained most in terms of LE and which country gained least?
```{r}
life_expectancy %>% 
  arrange(desc(growthLE)) %>% 
  head(1)
life_expectancy %>% 
  arrange(growthLE) %>% 
  head(1)
```
#### Maldives gained the most LE, Ukraine gained the least

### 5.2.4
How many countries were there where LE improved less than 5 years?
```{r}
life_expectancy %>% 
  filter(growthLE < 5) %>% 
  summarize(count = n())
```
### 5.2.5
How many countries were there were LE decreased over this time period?
```{r}
life_expectancy %>% 
  filter(growthLE < 0) %>% 
  summarize(count = n())
```
