---
title: "ps05"
output: html_document
date: "2023-11-19"
---
## 1.1
```{r}
library(tidyverse)
library(ggplot2)
library(nycflights13)

flights %>% 
  sample_n(5)
```

## 1.1.2.a
```{r}
print(sum(!(nchar(flights$origin) == 3 & nchar(flights$dest) == 3)))
```
check if there are airport codes that are not 3 characters long

## 1.1.2.b
```{r}
print(sum(grepl("\\d", flights$origin) & grepl("\\d", flights$dest)))
```
check if there are airport codes that contains digits

## 1.1.3
```{r}
summary(flights$arr_delay)
```
There are 9430 missing arrival delay in data. It is a problem because if we want to analyze delay time, missing a relatively large portion of arrival delay might result in bias of the result

## 1.1.4
based on the summary from above, the max delay is approximately 21 hours, which is plausible because airport has certain compensation for long time delay like that. It's also possible to have about one and a half hour early

## 1.1.5
```{r}
flights %>% 
  select(arr_delay, dest) %>% 
  group_by(dest) %>% 
  summarize(average_delay = mean(arr_delay, na.rm = TRUE)) %>% 
  arrange(desc(average_delay)) %>% 
  head(3)
```

## 1.1.6
```{r}
month_delay <- flights %>% 
  select(arr_delay, month) %>% 
  group_by(month) %>% 
  summarize(average_delay = mean(arr_delay, na.rm = TRUE))
head(month_delay, 12)
```

## 1.1.7
```{r}
ggplot(month_delay, aes(x = factor(month), y = average_delay)) +
  geom_bar(stat = "identity", aes(fill = factor(month))) +
  labs(x = "Month", y = "Average Delay", title = "Monthly Average Flight Delays")
```
There's not a clear pattern here, but from Septermbet to November, it experience a really low delay even arrive early. There's a relatively high delay during June, July, and December. It's probably due to holidays and increasing number of flights.

## 1.1.8
```{r}
daily_delay <- flights %>%
  filter(year == 2013) %>% 
  mutate(date = make_date(year, month, day),
         season = case_when(
           month %in% c(12, 1, 2) ~ "Winter",
           month %in% 3:5 ~ "Spring",
           month %in% 6:8 ~ "Summer",
           month %in% 9:11 ~ "Fall"
         )) %>%
  group_by(date, season) %>%
  summarize(average_delay = mean(arr_delay, na.rm = TRUE), .groups = 'drop')

ggplot(daily_delay, aes(x = date, y = average_delay, color = season)) +
  geom_point(alpha = 0.6) +
  scale_color_manual(values = c("Winter" = "blue", "Spring" = "green", "Summer" = "red", "Fall" = "orange")) +
  labs(x = "Date", y = "Average Delay", title = "Daily Average Flight Delays in 2013")
```
the delay is more common in summer compares to winter and the delay is more common in spring compare to fall

## 1.2.1
```{r}
chicago <- flights %>% 
  filter(dest == "ORD")
nrow(chicago)
```
there are 17283 flights to chicago

## 1.2.2
```{r}
chicago %>% 
  arrange(air_time) %>% 
  select(year, month, day, dep_time, air_time, carrier, origin) %>% 
  slice(1)
```
the fastest flight to chicago is 87 minutes from Newark Liberty International Airport by United Airline on 2013/08/21 at 16:04

## 1.2.3
The typical flight times for direct flights from NYC to ORD is 160 minutes, the data is only half of it do it doesn't seems to be realistic.

## 1.2.4
```{r}
chicago %>% 
  arrange(desc(air_time)) %>% 
  select(year, month, day, dep_time, air_time, carrier, origin) %>% 
  slice(1)
```
The slowest flight to chicago is 198 minutes from Newark Liberty International Airport by United Airline on 2013/08/21 at 06:38. The result does make sense since it's only a bit slower than average flight time.

## 1.2.5
```{r}
chicago %>%
  mutate(speed_mph = distance / (air_time / 60)) %>%
  filter(!is.na(speed_mph)) %>%
  arrange(speed_mph) %>%  
  select(time_hour, carrier, dep_delay, speed = speed_mph) %>% 
  slice(c(1:3, (n()-2):n())) 
```

## 1.2.6
```{r}
chicago %>%
  mutate(speed_mph = distance / (air_time / 60)) %>% 
  filter(!is.na(speed_mph) & !is.na(air_time)) %>%
  ggplot(aes(y = dep_delay, x = speed_mph)) +
  geom_point(alpha = 0.2) +
  labs(x = "Speed (mph)", y = "Departure Delay (minutes)", 
       title = "Departure delay vs speed for flights to Chicago (scatter)")
```
```{r}
chicago %>%
  mutate(speed_mph = distance / (air_time / 60)) %>% 
  filter(!is.na(speed_mph) & !is.na(air_time)) %>%
  ggplot(aes(y = dep_delay, x = speed_mph)) +
  geom_line(alpha = 0.2) +
  labs(x = "Speed (mph)", y = "Departure Delay (minutes)", 
       title = "Departure delay vs speed for flights to Chicago (line)")
```
I tried scatter plot because speed and departure delay are two nearly continous variable. I also tried to use a line plot but it doesn't show a smooth transition. On the scatter plot, I noticed that the relationship of speed and departure delay was kind of similar to normal distribution where points cluster around the middle range at around 380 mph. 

## 2.1.1
```{r}
gm <- read_delim("../Data/gapminder.csv.bz2")
dim(gm)
```
we have 25 columns and 13055 rows

## 2.1.2
```{r}
gm %>% 
  sample_n(5)
```
It looks ok because it contains all the varibales we want. It contains some NA values.

## 2.1.3
```{r}
gm %>% 
  distinct(iso3) %>%
  nrow()
gm %>% 
  distinct(iso2) %>%
  nrow()
gm %>% 
  distinct(name) %>%
  nrow()

```
there are 253 countries based on iso3, 249 countries based on iso2, 250 countries based on name

## 2.1.4.a
```{r}
gm %>%
  group_by(iso2) %>%
  summarize(unique_names = n_distinct(name)) %>%
  filter(unique_names > 1) %>% 
  head()

gm %>% 
  filter(is.na(iso2)) %>%
  distinct(name)
```
 A country with missing name and Namibia has a missing iso2

## 2.1.4.b
```{r}
gm %>%
  group_by(name) %>%
  summarize(unique_iso3 = n_distinct(iso3), .groups = 'drop') %>%
  filter(unique_iso3 > 1) %>% 
  head()

gm %>% 
  filter(is.na(name)) %>% 
  distinct(iso3)
```
Country has mroe than one iso3 code are CHANISL, GBM, KOS, NLD CURACAO

## 2.2.1
```{r}
selected <- c("United States of America", "China", "India", "Canada", "France")
selected_gm <- gm %>%
  filter(name %in% selected  & !is.na(co2))

ggplot(selected_gm, aes(x = time, y = co2, color = name)) +
  geom_line() +
  labs(title = "CO2 Emissions Changed Over Time",
       x = "Year",
       y = "Total CO2 Emissions",
       color = "Country")
```
China has significantly more CO2 emission over 2005 compares to other counties, Canada and France keep a low CO2 emission
```{r}
ggplot(selected_gm, aes(x = time, y = co2_PC, color = name)) +
  geom_line() +
  labs(title = "CO2 Emissions per Capita Changed Over Time",
       x = "Year",
       y = "Total CO2 Emissions Per Capita",
       color = "Country")
```
The figure suggests that US has a highest CO2 Emmisions per Capita and India has the lowest CO2 Emmisions per Capita

## 2.2.3
```{r}
avg_continent <- gm %>% 
  filter(!is.na(region)) %>% 
  group_by(region, time) %>% 
  summarize(average_CO2_per_capita = mean(co2_PC, na.rm = TRUE))
head(avg_continent)
```
## 2.2.4
```{r}
avg_continent %>%
  filter(time %in% c(1960, 2016)) %>% 
  ggplot(aes(x = region, y = average_CO2_per_capita, fill = as.factor(time))) +
  geom_bar(stat = "identity", 
           position = position_dodge(), 
           color = "gray20") +
  labs(title = "Average CO2 Emissions Per Capita Across 
       \nin Continents in 1960 and 2016",
       x = "Continent",
       y = "Average CO2 Emissions Per Capita",
       fill = "Year")
```

## 2.3.1
```{r}
letf1960 <- gm %>%
  filter(time == 1960 &
           !is.na(fertilityRate) & 
           !is.na(lifeExpectancy)& 
           !is.na(region))

ggplot(letf1960, aes(x = fertilityRate, 
           y = lifeExpectancy, 
           size = totalPopulation,
           color = region)) +
  geom_point(alpha = 0.7) +
  scale_size(range = c(1, 10), name = "Population") +
  scale_color_brewer(palette = "Set1") +
  labs(title = "Total Fertility vs Life Expectancy in 1960",
       x = "Total Fertility",
       y = "Life Expectancy",
       color = "Continent")
```
total fertility and life expectancy is generally inversely proportional
Europe has overall lowet fertility and highest life expectancy
Africa has overall highest fertility and lowest life expectancy

## 2.3.3
```{r}
letf2019 <- gm %>%
  filter(time == 2019 &
           !is.na(fertilityRate) & 
           !is.na(lifeExpectancy)& 
           !is.na(region))

ggplot(letf2019, aes(x = fertilityRate, 
           y = lifeExpectancy, 
           size = totalPopulation,
           color = region)) +
  geom_point(alpha = 0.7) +
  scale_size(range = c(1, 10), name = "Population") +
  scale_color_brewer(palette = "Set1") +
  labs(title = "Total Fertility vs Life Expectancy in 1960",
       x = "Total Fertility",
       y = "Life Expectancy",
       color = "Continent")
```
the world's lfe expectancy increased over the past 60 years
and the total fertility decreased
The total fertility of Americas and Asia decreased significantly

## 2.3.4
```{r}
growLE <- gm %>%
  filter(time %in% c(1960, 2019) &
           !is.na(fertilityRate) & 
           !is.na(lifeExpectancy)& 
           !is.na(region)) %>%
  group_by(region, time) %>%
  summarize(averageLE = mean(lifeExpectancy, na.rm = TRUE))
head(growLE, 12)
```
## 2.3.5
```{r}
growLE %>%
  mutate(LE_growth = averageLE - lag(averageLE)) %>%
  filter(!is.na(LE_growth)) %>%
  arrange(desc(LE_growth))
```
Asia experienced highest average LE growth, and Eruope experienced lowest average LE growth
## 2.3.6
```{r}
gm %>% 
  filter(name == "United States of America" | 
           name == "Sweden" |
           name == "China" | 
           name == "Gabon" |
           name == "Mongolia" |
           name == "Turkey"&
           !is.na(fertilityRate) & 
           !is.na(lifeExpectancy)) %>% 
  ggplot(aes(x = fertilityRate, 
             y = lifeExpectancy, 
             color = name)) +
  geom_path() +
  geom_point(size = 1, alpha = 0.5) +
  labs(title = "Total Fertility vs Life Expectancy",
       x = "Total completed fertility",
       y = "Life expectancy at birth",
       color = "Country")
```

## 2.3.7
```{r}
gm%>%
  filter(!is.na(name), time == "1960" & !is.na(lifeExpectancy)) %>%
  mutate(rank1960 = rank(desc(lifeExpectancy))) %>%
  filter(name == "United States of America")%>%
  filter(!is.na(rank1960)) %>%
  pull(rank1960) %>%
  cat("The rank of US in terms of life expectancy in 1960 is", ., "\n")

gm%>%
  filter(!is.na(name), time == "2019" & !is.na(lifeExpectancy)) %>%
  mutate(rank2019 = rank(desc(lifeExpectancy))) %>%
  filter(name == "United States of America")%>%
  filter(!is.na(rank2019)) %>%
  pull(rank2019) %>%
  cat("The rank of US in terms of life expectancy in 2019 is", ., "\n")
```

## How many hours
```{r}
cat("more than I expected :(")
```

