---
title: "lab7"
output: html_document
date: "2023-11-16"
---
## 1.1
Filter only those people with income > 0
```{r}
library(tidyverse)
library(ggplot2)
ih <- read.delim("../Data/income-height.csv.bz2")
sample_n(ih, 5)

ih <- ih %>% 
  filter(income > 0)
```

## 1.2
For each person in the dataset, compute 5-inch height category.
```{r}
ih <- ih %>% 
  mutate(height_category =(height %/% 5) * 5)

head(ih)
```
## 1.3
Make a barplot for average income vs each age category
```{r}
ih %>% 
  group_by(height_category) %>% 
  summarize(average_income = mean(income, na.rm = TRUE)) %>% 
  ggplot(aes(height_category, average_income,
             fill = factor(height_category))) + 
  geom_col() + 
  labs(title = "Average Income for Each Age Category",
       x = "Height Category (inches) ",
       y = "Average Yearly Income ($)",
       fill = "Height Category")
```
### the plot suggest that there is a general trend about taller people higher income, except people in 75 inches genrally earn less than people in 70 inches. People in 80 inches earn a lot more than others

## 1.4
Make a barplot with two sets of bars (for men/women)
next to each other, filled with a different colors.
```{r}
ih %>% 
  group_by(height_category, sex) %>% 
  summarize(average_income = mean(income, na.rm = TRUE)) %>% 
  ggplot(aes(height_category, average_income,
             fill = sex)) + 
  geom_col(position = "dodge") + 
  labs(title = "Average Income for Each Age Category",
       x = "Height Category (inches) ",
       y = "Average Yearly Income ($)",
       fill = "Height Category")
```
## 1.5
```{r}
ih %>% 
  ggplot(aes(factor(height_category), income)) + 
  geom_boxplot(aes(fill = height_category)) + 
  labs(title = "Income for Each Age Category",
       x = "Height Category (inches) ",
       y = "Income ($)")
```

## 1.6
```{r}
ih %>% 
  ggplot(aes(factor(height_category), income, fill = sex)) +
    geom_boxplot() +
    labs(title = "Income by Height Category(Seperate Gender)",
         x = "Height Category",
         y = "Income")
```

### men generally earn a little more than women with same height category，There is a general trend for taller people has more income. Men with 80 inches earn significantly more.

## 2.1
```{r}
covid <- read_delim("../Data/covid-scandinavia-wide.csv.bz2")

covid %>% 
  sample_n(5)
```

### a row represent the confirmed cases and death cases in a certain date in countries in Scandinavia (Denmark, Finland, Norway and Sweden)

## 2.2
```{r}
first_date <- min(covid$date, na.rm = TRUE)
last_date <- max(covid$date, na.rm = TRUE)
print(paste("First date in data is", first_date, "and last date in data is", last_date))
```

## 2.3
```{r}
covid %>% 
  filter(country == "Denmark") %>% 
  ggplot(aes(Confirmed, Deaths, color = "RED")) + 
  geom_point(alpha = 1/5, size = 1) +
  labs(title = "Covid Confirmed Cases vs Deaths cases in Denmark",
       x = "Confirmed Cases", 
       y = "Deaths Cases")
```

## 2.4
```{r}
covid %>% 
  ggplot(aes(Confirmed, Deaths),
         col = country) + 
  geom_point(alpha = 1/5, size = 1, aes(color = country)) +
  labs(title = "Confirmed Cases vs Deaths cases in Denmark",
       x = "Confirmed Cases", 
       y = "Deaths Cases")
```

## 2.5
```{r}
covid %>% 
  ggplot(aes(Confirmed, Deaths),
         col = country) + 
  geom_path(aes(color = country)) +
  labs(title = "Confirmed Cases vs Deaths cases in Denmark",
       x = "Confirmed Cases", 
       y = "Deaths Cases")
```