---
title: "INFO370lab04"
author: "Marco"
date: "`r Sys.Date()`"
output: html_document
---

```{r}
library(tidyverse)
```


## 1
### 1.1
```{r}
temp <- read_delim("../../data/UAH-lower-troposphere-wide.csv.bz2")
dim(temp)
sample_n(temp, 5)
```
### 1.2
```{r}
range(temp$year)
range(temp$month)
```
Year range from 1978 to 2024, month range from 1 to 12

### 1.3
```{r}
temp <- temp %>% 
  mutate(time = year + month/12)

head(temp)
```
### 1.4
```{r}
temp %>% 
  ggplot(aes(time, globe)) + 
  geom_point(color = "gold2", alpha = 0.7) +
  geom_smooth(method = "lm", se = FALSE, color = "purple")
```
### 1.5
We are not capture the trend within year, there might be some variation between that. But we would focus on the overall trend so it is a good approximation.

### 1.6
```{r}
model <- lm(globe ~ time, data = temp)
summary(model)
```

### 1.7
The number means that as year increase time(year) by one, the global temperature deviation would increase by a mean of 0.01521. It is statistically significant at 0 confidence level.


### 1.8
Intercept means that when time 2000 years ago. The global temperature deviation has a mean of  -30.48959.

### 1.9
The “GTR” (global temperature report) from University of Alabama, Huntsville also shows a overall increasing trend.

## 2
### 2.1
```{r}
gdp <- read_delim("../../data/gapminder.csv.bz2")
dim(gdp)
sample_n(gdp, 5)
```

### 2.2
```{r}
gdp <- gdp %>% 
  filter(time == 2016) %>% 
  select(time, GDP_PC, lifeExpectancy)
head(gdp)
```

### 2.3
```{r}
model2 <- lm(lifeExpectancy ~ GDP_PC, data = gdp)
summary(model2)
```

### 2.4
Intercept means that country with 0 GDP per capita ($) would have a mean of 6.864e+01 (years) Life expectancy at birth
The GDP_PC means that as GDP per capita increase 1 unit, the life expectancy would increase by 2.477e-04 (years)
The relationship between GDP and life expectancy statistically significant, since p value is less than 0.5


### 2.5
```{r}
model2 <- lm(lifeExpectancy ~ I(GDP_PC/1000), data = gdp)
summary(model2)
```

### 2.6
The intercept keep the same. The slope would be as gdp increase by 1000$ the life expectancy would increase by 0.24766(year). It would make explaination easier as we want to study life expectancy vs month

### 2.7
```{r}
gdp %>% 
  ggplot(aes(GDP_PC, lifeExpectancy)) + 
  geom_point(color = "gold2", alpha = 0.7) +
  geom_smooth(method = "lm", se = FALSE, color = "purple")
```
The relationship is not really linear.

```{r}

model2 <- lm(lifeExpectancy ~ log(GDP_PC/1000), data = gdp)
summary(model2)

gdp %>% 
  ggplot(aes(log(GDP_PC/1000), lifeExpectancy)) + 
  geom_point(color = "gold2", alpha = 0.7) +
  geom_smooth(method = "lm", se = FALSE, color = "purple") 
```