---
title: "Info370_Lab03"
author: "Marco"
date: "`r Sys.Date()`"
output: html_document
---
## 1
### 1.1
$$
E(x) = \sum_{i=1} ^n x_1 \times p_1
$$

```{r}
Ex <- 1*0.5 + 0*0.5
Ex
```

### 1.2

$$
Var(x) = E[(x - \mu)^2]\\
Var(x) = E[x^2] - \mu^2
$$

```{r}
varx <- (0^2 * 0.5 + 1^2 * 0.5) - (0.5)^2
varx
```

### 1.3
```{r}
S <- 10
n <- 1
p <- 0.5
X <- rbinom(S, n, p)
mean(X)
var(X)
```

### 1.4
```{r}
S <- 100000
n <- 1
p <- 0.5
X <- rbinom(S, n, p)
mean(X)
var(X)
```

### 1.5
The sample mean and variance are pretty close to theoretical value as sample size increase to 100000

## 2
### 2.1
```{r}
Ex <- 0*0.16 + 1 * 0.48 + 2*0.36
Ex
```

### 2.2
```{r}
Varx <- (0^2 * 0.16 + 1^2 * 0.48 + 2^2 * 0.36) - (1.2)^2
Varx
```

### 2.3
```{r}
S <- 10
n <- 2
p <- 0.6
X <- rbinom(S, n, p)
mean(X)
var(X)
```

### 2.4
```{r}
S <- 100000
n <- 2
p <- 0.6
X <- rbinom(S, n, p)
mean(X)
var(X)
```

### 2.5
The sample mean and variance are pretty close to the theoretical value as sample size increased to 100000

## 3
### 3.1
```{r}
CLT <- function(S, sigma, R) {

  averages <- numeric(R)
  
  for (i in 1:R) {
    samples <- rlnorm(S, 0, sigma)
    averages[i] <- mean(samples)
  }
  
  hist(averages, breaks = 30, main = paste("Histogram of Averages (S =", S, ", sigma =", sigma, ")"),
       xlab = "Averages", col = "pink")
}
```

### 3.3
```{r}
CLT(20, 0.2, 10000)
```

### 3.4
```{r}
CLT(1, 1, 10000)
```

### 3.5
```{r}
CLT(20, 1, 10000)
```

### 3.6
```{r}
CLT(1, 2, 10000)
```

### 3.7
```{r}
CLT(20, 2, 10000)
```

### 3.8 
When S = 1, the standard deviation = 2 case is more right skewed


### 3.9 
When S = 20, the standard deviation = 0.2 looks more normal.

When standard deviation is fixed, more sample tends to make the distribution more normal
When sample size is fixed, less standard deviation make function more normal
The larger sample size the more normal it looks. While larger sd less normal it looks.

After testing I notice that standard deviation affects skewness more than sample size.
If the sd is really small such as 0.1, sample size won't change much of how normal it looks